---
title: "run.Rmd"
author: "BCM MSPC"
date: '`r Sys.Date()`'
output:
  rmdformats::downcute:
    downcute_theme: "chaos"
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    code_folding: "hide"
    css: "custom.css"
params:
  volcanodir: "/mnt/e/MSPC001161/results/MSPC001161_20240815/volcano/"
  rankfiledir: "/mnt/e/MSPC001161/results/tackle2_linregres/rnkfiles/"
  #gct_path: "../results/MSPC001128/export/MSPC001128_data_gct_all_1.0more_nonzero_normtype_median_n6x2995.gct"
  gct_path: "/mnt/e/MSPC001161/results/MSPC001161_20240815/export/MSPC001161_20240815_data_gct_all_0.66more_nonzero_normtype_median_n13x3317.gct"
  cut_by: none
  # ranks_from: "volcano"  # choices are volcano and gct
  ranks_from: "volcano"  # choices are volcano and gct
  species: "Rattus norvegicus"	
  savedir: "/mnt/e/MSPC001161/results/tackle2_linregres"
  genesets_json: >
    [
      {"category": "H", "subcategory": "", "collapse" : "FALSE"},
      {"category": "C2", "subcategory": "CP:KEGG", "collapse": "TRUE"},
      {"category": "C2", "subcategory": "CP:REACTOME", "collapse": "TRUE"},
      {"category": "C5", "subcategory": "GO:MF", "collapse" : "TRUE"},
      {"category": "C5", "subcategory": "GO:BP", "collapse" : "TRUE"},
      {"category": "C5", "subcategory": "GO:CC", "collapse" : "TRUE"} ,
      {"category": "C5", "subcategory": "All", "collapse" : "TRUE"}
    ]

---

# setup

```{r setup, include=TRUE}
suppressPackageStartupMessages(library(purrr)) # for map()
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(fs)) # for dir_ls()
suppressPackageStartupMessages(library(fgsea))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(msigdbr))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(tictoc))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(reactable))
suppressPackageStartupMessages(library(ggalt))
suppressPackageStartupMessages(library(rmdformats))
# suppressPackageStartupMessages(library(logger)) # logging never seems to work right inside Rmd..
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
options(reactable.theme = reactableTheme(
  color = "hsl(233, 9%, 87%)",
  backgroundColor = "hsl(233, 9%, 19%)",
  borderColor = "hsl(56, 100%, 58%, 0.76)",
  stripedColor = "hsl(233, 12%, 22%)",
  highlightColor = "hsl(233, 12%, 24%)",
  inputStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  selectStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonHoverStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonActiveStyle = list(backgroundColor = "hsl(233, 9%, 28%)")
))
```

```{r}
library(future)
n_cores <- future::availableCores() - 1
future::plan(future::multisession, workers = min(n_cores, 8))

io_tools <- new.env()
source("./R/io.R", local = io_tools)

util_tools <- new.env()
source("./R/utils.R", local = util_tools)

geneset_tools <- new.env()
source("./R/geneset_utils.R", local = geneset_tools)

fgsea_tools <- new.env()
source("./R/fgsea.R", fgsea_tools)

plot_tools <- new.env()
source("./R/plot.R", local = plot_tools)

plot_utils <- new.env()
source("./R/plot_utils.R", local = plot_utils)

pca_tools <- new.env()
source("./R/pca.R", local = pca_tools)

savedir <- params$savedir
if (is.null(savedir)) {
  savedir <- "./plots"
}

save_func <- util_tools$make_partial(
  plot_utils$plot_and_save,
  path = savedir
)


log_msg <- util_tools$make_partial(util_tools$log_msg)
# log_message <- make_partial(util_tools$log_message)
log_msg(msg = paste0("===\n*starting tackle2*\n==="))

TESTING <- FALSE
```
# explanations

```{r, label=test, eval=F, include=F}
datas1 <- purrr::map(1:3, ~ fgsea_tools$simulate_preranked_data(seed = 4321))
datas1 %<>% purrr::map(~ .x %>% mutate(value = value + rnorm(nrow(.), sd = .1)))
datas2 <- purrr::map(1:3, ~ fgsea_tools$simulate_preranked_data(seed = 1234))
datas2 %<>% purrr::map(~ .x %>% mutate(value = value + rnorm(nrow(.), sd = .1)))
datas3 <- purrr::map(1:3, ~ fgsea_tools$simulate_preranked_data(seed = 9999))
datas3 %<>% purrr::map(~ .x %>% mutate(value = value + rnorm(nrow(.), sd = .1)))
rnkdfs <-
  c(
    datas1,
    datas2,
    datas3
  )
names(rnkdfs) <- paste0("test", 1:length(rnkdfs))
# rankobj <- io_tools$ranks_dfs_to_lists(list(test = datas[[1]]))
# .res <- rankobj[[1]] %>% fgsea_tools$run_one(geneset_list)
# .name <- .res[1, "pathway"][[1]]

ranks_list <- rnkdfs %>% io_tools$ranks_dfs_to_lists()
rankobjs <- io_tools$ranks_dfs_to_lists(rnkdfs = rnkdfs)

# enplot_data <- fgsea::plotEnrichmentData(geneset_list[[.name]], rankobj[[1]])
# .rnkorder <- -rankobj[[1]] %>% (rank)
# rankobj[[1]]['rank'] <- .rnkorder

# geneset <- geneset_tools$get_collection("H", "")
# geneset_list <- geneset_tools$genesets_df_to_list(geneset)

genesets_of_interest <- tibble::tribble(
  ~category, ~subcategory, ~collapse,
  "H", "", FALSE,
  "C5", "GO:MF", TRUE
) %>% mutate(collection_name = stringr::str_c(category, subcategory, sep = "_")) # todo put this somewhere else
list_of_geneset_dfs <- geneset_tools$get_collections(genesets_of_interest)
genesets_list_of_lists <- list_of_geneset_dfs %>% purrr::map(geneset_tools$genesets_df_to_list)

.results_list <- fgsea_tools$run_all_pathways(
  geneset_lists = genesets_list_of_lists,
  ranks = rankobjs,
  genesets_additional_info = genesets_of_interest, # dataframe with collapse indicator
  parallel = T
)
results_list <- .results_list

all_gsea_results <- fgsea_tools$concat_results_all_collections(results_list)
# named_data = list(data=data)


.datas <- rnkdfs %>% purrr::imap(~ {
  .x %>% dplyr::rename(!!.y := value)
})
# .datas[[1]] %>% head()
.mat <- base::Reduce(
  accumulate = F,
  f = function(...) full_join(..., by = "id"),
  x = .datas
) %>% as.data.frame()

rownames(.mat) <- .mat$id
.mat$id <- NULL
.meta <- data.frame(
  id = colnames(.mat),
  group = c(rep("A", 3), rep("B", 3), rep("C", 3))
)
rownames(.meta) <- .meta$id
.rdesc <- data.frame(id = rownames(.mat), dummy = "X")
rownames(.rdesc) <- rownames(.mat)
gct <- new("GCT",
  mat = .mat %>% as.matrix(),
  cdesc = .meta,
  rdesc = .rdesc
)

TESTING <- TRUE
```

## run these to examine what genesets are available from msigdbr


```{r, label='show species', eval=F, include=T}
msigdbr::msigdbr_species()
```

```{r, label='show collections', eval=F, include=T}
msigdbr::msigdbr_collections()
```

# execution

## load gene sets

```{r, label='load params'}
# genesets_of_interest <- jsonlite::parse_json(params$genesets_json)
if (!exists("genesets_of_interest")) {
  json_str <- params$genesets_json
  genesets_of_interest <- io_tools$load_genesets_from_json(json_str)
}


# SPECIES <- "Homo sapiens" # or "Mus musculus" or Saccharomyces cerevisiae, and others
# genesets_of_interest %>% purrr::map(~{print(.x$category); print(.x$subcategory)})
```


```{r, label='get genesets'}
species <- params$species
if (is.null(species)) species <- "Homo sapiens"

if (!exists("list_of_geneset_dfs")) {
  list_of_geneset_dfs <- genesets_of_interest %>%
    geneset_tools$get_collections(species = species)
}

if (!exists("genesets_list_of_lists")) {
  genesets_list_of_lists <- list_of_geneset_dfs %>% purrr::map(geneset_tools$genesets_df_to_list)
}
```

## prepare ranks

```{r, label='locate ranks'}
rankfiledir <- params$rankfiledir
volcanodir <- params$volcanodir
gct_path <- params$gct_path
ranks_from <- params$ranks_from

log_msg(paste0("rankfiledir: ", rankfiledir))
log_msg(paste0("volcanodir: ", volcanodir))
log_msg(paste0("gct_path: ", gct_path))
log_msg(paste0("ranks from: ", ranks_from))


if (!exists("TESTING")) TESTING <- FALSE


if ((!is.null(volcanodir) && any(!file.exists(volcanodir))) &&
  (!is.null(gct_path) && !file.exists(gct_path)) &&
  (!is.null(rankfiledir) && !file.exists(rankfiledir)) &&
  (!TESTING == TRUE)) {
  print("no files present")
  log_msg(msg = "no files present")
  knitr::knit_exit()
}


log_msg(msg = paste0("TESTING: ", TESTING))
```

```{r, label='load ranks'}
load_and_process_ranks <- function(params) {
  rankfiledir <- params$rankfiledir
  volcanodir <- params$volcanodir
  gct_path <- params$gct_path
  ranks_from <- params$ranks_from

  log_msg(msg = paste0("ranks from : ", ranks_from))
  log_msg(msg = paste0("rankfiledir : ", rankfiledir))

  if (!is.null(rankfiledir) && file.exists(rankfiledir)) { # 
    rnkfiles <- dir_ls(path = rankfiledir, regexp = ".*\\.rnk$", fail = FALSE)
    log_msg(msg = paste0("looking for rank files in ", rankfiledir))
    if (length(rnkfiles) > 0) {
      log_msg(msg = paste0("found ", length(rnkfiles), " rankfiles"))
      rnkdfs <- rnkfiles %>% io_tools$load_rnkfiles()
      names(rnkdfs) <- names(rnkdfs) %>%
        fs::path_file() %>%
        fs::path_ext_remove()
      ranks_list <- rnkdfs %>% io_tools$ranks_dfs_to_lists()
      return(ranks_list)
    } # exit and we're done
    log_msg(msg = "couldn't find any rnkfiles")
    if (ranks_from == "volcano") {
      if (is.null(volcanodir) || length(volcanodir) == 0 || any(!file.exists(volcanodir))) {
        stop("improper volcanodir specification")
      }
      log_msg(msg="saving rankfiles from volcano output. using signedlogp as value")
      rnkdfs <- io_tools$create_rnkfiles_from_volcano(volcanodir, value_col = "signedlogP")
      rnkdfs %>% io_tools$write_rnkfiles(dir = rankfiledir) # and save
      names(rnkdfs) <- names(rnkdfs) %>%
        fs::path_file() %>%
        fs::path_ext_remove()
      log_msg(paste0("length of retrieved rankfiles: ", length(rnkdfs)))
      ranks_list <- rnkdfs %>% io_tools$ranks_dfs_to_lists()
      return(ranks_list)
    }
    if (ranks_from == "gct" && !is.null(gct_path)) {
      rnkdfs <- io_tools$create_rnkfiles_from_emat(gct_path, apply_z_score = TRUE)
      names(rnkdfs) <- names(rnkdfs) %>%
        fs::path_file() %>%
        fs::path_ext_remove()
      log_msg(msg=paste0("length of retrieved rankfiles: ", length(rnkdfs)))
      ranks_list <- rnkdfs %>% io_tools$ranks_dfs_to_lists()
    }
    # not sure if this level of flow is relevant, refactor later
    rnkdfs %>% io_tools$write_rnkfiles(dir = rankfiledir)
    return(ranks_list)
  }
  # # Always reload rank files from new location to ensure proper format
  # log_msg(paste0('rankfiledir: ', rankfiledir))
  # rnkfiles <- dir_ls(path=rankfiledir, regexp = ".*\\.rnk$", fail = FALSE)
  # log_msg(paste0('n rank files: ', length(rnkfiles)))
  # log_msg(paste0('rank files: ', str_c(rnkfiles, collapse='\n')))
  #
  # rnkdfs <- rnkfiles %>% io_tools$load_rnkfiles()
  # names(rnkdfs) <- names(rnkdfs) %>% path_file() %>% path_ext_remove()
  # ranks_list <- rnkdfs %>% io_tools$ranks_dfs_to_lists()
  #
  # log_msg(paste0('names ranks list: ', str_c(names(ranks_list), collapse='\n')))
  # return(ranks_list)
}
ranks_list <- load_and_process_ranks(params)

```


```{r, label='maybe load gct', eval=!TESTING}
# this whole section can (and should) be skipped if TESTING == TRUE


if (!exists("gct") || is.null(gct)) {
  if (!is.null(gct_path) && file.exists(gct_path)) {
    log_msg(msg = paste0("reading gct file: ", gct_path))
    gct <- cmapR::parse_gctx(gct_path)
  } else {
    gct <- NULL
  }
}
# log_msg(msg = paste0("gct: ", class(gct)))

# rnkfiles <- dir_ls(path=rankfiledir, regexp = ".*rnk$", fail = F)
# log_msg(msg=paste0('n rank files: ', length(rnkfiles)))
# log_msg(msg=paste0('rank files: ', str_c(rnkfiles, sep='\n')))
# rnkdfs <- rnkfiles %>% io_tools$load_rnkfiles()
# ranks_list <- rnkdfs %>% io_tools$ranks_dfs_to_lists()

log_msg(msg = paste0("names ranks list "))
log_msg(msg = str_c(names(ranks_list), collapse = "\n"))
```


# run fgsea
```{r, label='run fgsea', cache=T}
(names(ranks_list))
(names(genesets_list_of_lists))
if (TESTING == FALSE) {
  results_list <- fgsea_tools$run_all_pathways(genesets_list_of_lists,
    ranks_list,
    parallel = T,
    genesets_additional_info = genesets_of_interest
  ) # parallel if you c
}
# results_list2 <- .GlobalEnv$run_all_pathways(geneset_lists, ranks_list, parallel=T) # parallel if you c

log_msg(msg = "names gsea results list: ")
log_msg(msg = str_c(names(results_list), sep = "\n"))

log_msg(msg = "comparison  names gsea results list: ")
log_msg(msg = str_c(names(results_list[[1]]), sep = "\n"))
```



## explore the results

results list contains a named list of pathways

```{r}
names(results_list)
```

which contain a named list of statistics

```{r}
names(results_list[[1]])
```

```{r}
results_list[[1]][[1]] %>%
  dplyr::select(-c("leadingEdge")) %>%
  arrange(pval) %>%
  head(30) %>%
  reactable(
    resizable = T,
    wrap = F,
    bordered = T,
    filterable = F,
    searchable = T,
    defaultPageSize = 6,
    striped = T,
    compact = T,
    showSortable = T,
    showPageSizeOptions = T,
    columns = list(
      pval = colDef(format = colFormat(digits = 2)),
      padj = colDef(format = colFormat(digits = 2)),
      log2err = colDef(format = colFormat(digits = 2)),
      ES = colDef(format = colFormat(digits = 4)),
      NES = colDef(format = colFormat(digits = 4))
    )
  )
```



```{r, label = "save gsea result tables", eval=T}
._ <- results_list %>%
  io_tools$save_gsea_results(savedir = file.path(savedir, "gsea_tables"))
```


## barplots


```{r, label='one barplot'}
results_list[[1]][[1]] %>%
  arrange(pval) %>%
  head(20) %>%
  plot_tools$barplot_with_numbers()
```
## plotting all barplots for each comparison within one geneset
```{r, label='all barplots, separate', eval=F}
results_list[[1]] %>% imap(
  ~ {
    sel <- .x %>%
      arrange(pval) %>%
      head(20)
    .title <- .y %>%
      fs::path_file() %>%
      fs::path_ext_remove() # %>% gsub(pattern="_", replacement=" ", x=.)
    plot_tools$barplot_with_numbers(sel, title = .title, use_custom_labeller = T)
  }
)
```

## plotting all barplots for each comparison across all genesets
```{r, label='all barplots, separate, save', fig.width=10, fig.height=8, eval=T}
log_msg(msg = "plotting individual barplots")
plts <- results_list %>% plot_tools$all_barplots_with_numbers(save_func = util_tools$make_partial(save_func, width = 12, height = 9))

# if you want to print the plots:
# for (plt in plts){
#   print(plt)
# }
```

## plot facetted barplot across genesets

prewritten function to merge gsea results across comparisons and make faceted barplot
```{r, label="combined barplots, save", fig.width=12, fig.height=12, eval=F}
log_msg(msg = "plotting faceted barplots")
plts <- results_list %>%
  plot_tools$process_results_across_rnks(
    save_func = util_tools$make_partial(save_func, width = 14, height = 18)
  )

for (plt in plts) {
  print(plt)
}


```



```{r, label='combine gsea matrices'}
log_msg(msg = "combining gsea marices")
all_gsea_results <- fgsea_tools$concat_results_all_collections(results_list)
```

# es plots
```{r, eval=F}

._ <- plot_tools$plot_top_ES_across(all_gsea_results,
                       ranks_list = ranks_list,
                       genesets_list_of_lists,
                       limit=30,
                       save_func = save_func
                           )

# this does one:
# plot_tools$plot_top_ES(all_gsea_results$`C5_GO:BP`,
#                        ranks_list = ranks_list,
#                        genesets_list_of_lists[["C5_GO:BP"]],
#                        limit=2,
#                        save_func = save_func 
#                        )


```


## heatmaps of results


```{r, label='plot one gsea level heatmap', fig.height=7, eval=T}
# below are the args for plotting gsea level heatmaps

log_msg(msg = "drawing a heatmap. save_func is null")

ht <- all_gsea_results$`C5_GO:CC` %>% plot_tools$plot_results_one_collection(
  metadata = NULL,
  cut_by = NULL,
  limit = 120,
  title = "<title>",
  main_pathway_ratio = .1,
  save_func = NULL
)

log_msg(msg = paste0("drew a heatmap, got this: ", class(ht)))
# log_msg(msg='calling ComplexHeatmap::draw')
# ComplexHeatmap::draw(ht,
#   heatmap_legend_side = "bottom",
#   padding = unit(c(2, 24, 2, 24), "mm"), # top, left, bottom, right
# )
ht
```

```{r, label='plot all gsea level heatmaps', fig.height=9, fig.width=11 }
log_msg(msg = "drawing all heatmap. save_func is not null")
# here is where changes to the save func are made (e.g. width, height, file format)

hts <- all_gsea_results %>% plot_tools$plot_results_all_collections(
  # limit=20,
  pstat_cutoff = 1,
  limit = 40,
  # save_func = .gsea_heatmap_save_func,
  save_func = save_func,
  cluster_columns=T,
  cluster_rows=T,
)

for (ht in hts) {
  draw(ht,
    heatmap_legend_side = "bottom",
    padding = unit(c(2, 24, 2, 24), "mm"), # top, left, bottom, right
  )
}


draw(ht,
  heatmap_legend_side = "bottom",
)
```


```{r, fig.height=12, label="plot edges"}
# plot_tools$make_heatmap_fromgct(
#   gct %>% subset_gct(rid=gct@rid %>% head(10))
# )

log_msg(msg = paste0("maybe plot edges"))
gct@cdesc %>% count(group)
gct@cdesc %<>% mutate(group = factor(group, levels = c("Pre_V", "Pre_A", "Control", "PCL", "PCL_Bi", "PCL_Bi_MSCs"), ordered = T))

if (!is.null(gct)) {
  ht_edge_plots <- plot_tools$plot_heatmap_of_edges(gct, results_list)
}
```

## pca and clustering


### pca

run pca
```{r, label='do pca'}
log_msg(msg = paste0("pca on gsea"))

if (exists("gct") && !is.null(gct)) {
  metadata <- gct@cdesc
} else {
  metadata <- NULL
}

pca_objects <- all_gsea_results %>% pca_tools$do_all(
  metadata = metadata
)
```


```{r, label='plot one level biplot'}
log_msg(msg = paste0("plot pca on biplot"))

pca_objects[[1]] %>% pca_tools$plot_biplot(
  top_pc = 3,
  showLoadings = T,
  labSize = 2,
  pointSize = 3,
  sizeLoadingsNames = 2,
  colby = "group", # or a string like 'group'. should not crash if not in metadata
  shape = NULL, # or a string like 'group'
  # encircle = ifelse(!is.null(colby), T, F),
  title = "first pca object",
  save_func = save_func
)
```


```{r, label='plot pca all biplots'}
log_msg(msg = paste0("plot pca all biplots"))
pca_objects %>% pca_tools$plot_all_biplots(
  save_func = save_func,
  top_pc = 3,
  showLoadings = T,
  labSize = 2,
  pointSize = 3,
  sizeLoadingsNames = 2,
  # colby = NULL,
)
```


## plot large tabular plots

this is slow
```{r, fig.width=10, fig.height=12, eval=F}
res <- plot_tools$plot_tables(
  results_list = results_list,
  ranks_list = ranks_list,
  pathways_list = genesets_list_of_lists["H_"]
)
```




```{r, label='..', eval=F,  include=F}
gct_path <- gct_path
gct <- cmapR::parse_gctx(gct_path)
all_gsea_results <- fgesa_tools$concat_results_all_collections(results_list)
.metadata <- gct@cdesc
.name <- "C5_GO:BP"


.pathways_withmeta <- all_gsea_results[[.name]] %>%
  left_join(gct@cdesc, by = c("var" = "id"))
.pathways_withmeta %<>% group_by(Model) %>%
  mutate(n_main = sum(mainpathway == T)) %>%
  mutate(ratio_main = n_main / n()) %>%
  ungroup()
# group_by(pathway) %>% summarize(n=sum(mainpathway==T)) %>% filter(n/.ntotal > .1) %>% distinct(pathway, .keep_all = T)

.ntotal <- all_res[[.name]]$var %>%
  unique() %>%
  length()
.pathways_filtered <- all_res[[.name]] %>%
  group_by(pathway) %>%
  summarize(n = sum(mainpathway == T)) %>%
  filter(n / .ntotal > .1) %>%
  distinct(pathway, .keep_all = T)
plt <- all_gsea_results[[.name]] %>%
  dplyr::filter(pathway %in% .pathways_filtered$pathway) %>%
  plot_tools$plot_results_one_collection(., metadata = .metadata, cut_by = "Model", title = .name)
```


## curve exploration

```{r, eval=F, include=F}
# Adjusting the formula with guessed parameters for a clearer sigmoid fit
# Example: a simple sigmoid with steepness and midpoint adjustments
# Ensure that you replace `.res2` and `pval` according to your actual data frame and variable names if they differ
sig_approx <- function(x) (-1 / (1 + exp(-4.1 * (x - 1.0))) + 1)

ggplot(.res2, aes(x = abs(NES), y = pval)) +
  geom_point() +
  stat_function(
    fun = sig_approx, # Adjust -0.1 (steepness) and 5 (midpoint) as needed
    color = "blue"
  ) +
  ggtitle("Fitting a Sigmoid Function")
```
```{r, eval=F}
sig_approx <- function(x) (-1 / (1 + exp(-4.1 * (x - 1.0))) + 1)
sig_approx2 <- function(x) 1 / sig_approx
sig_approx2 <- function(x) (1 + exp(-1.1 * (x - 1.0)) - 1)

ggplot(.res2, aes(y = abs(NES), x = pval)) +
  geom_point() +
  stat_function(
    fun = sig_approx2, # Adjust -0.1 (steepness) and 5 (midpoint) as needed
    color = "blue"
  ) +
  ggtitle("Fitting a Sigmoid Function")
```



## examine rankorder
not functional, need to move to supporting R file

```{r, eval=F}
# rankorder <- .GlobalEnv$get_rankorder(ranks_list[[1]], geneset_lists[[1]][[1]]  )



genesets_of_interest <- genesets_list_of_lists[[1]]
rankobject <- ranks_list[[1]]
# enplot_data <- plotEnrichmentData(genesets_of_interest[[1]], rankobject)
collection_name <- names(genesets_list_of_lists)[[1]]
geneset_of_interest_name <- names(genesets_of_interest)[[1]]

# enplot_raw <- plotEnrichment(genesets_of_interest[[1]], rankobject)
# enplot_data <- plotEnrichmentData(genesets_of_interest[[1]], rankobject)

rankorder_object <- fgsea_tools$get_rankorder(
  genesets_of_interest[[1]], rankobject,
  list_of_geneset_dfs[[collection_name]]
)

rankorder_object %>% plot_tools$plotES() +
  labs(title = geneset_of_interest_name)
```

```{r, eval=F}
make_selection <- function(x) {
  x <- x %>% arrange(rank)
  es_min <- x %>%
    pull(ES) %>%
    min(na.rm = T)
  es_max <- x %>%
    pull(ES) %>%
    max(na.rm = T)
  .first <- which(x$ES == es_min)
  .second <- which(x$ES == es_max)
  .bot <- x %>% tail(dim(x)[1] - .first)
  .top <- x %>% head(.second)
  browser()
  return(bind_rows(.top, .bot))
}

with(
  rankorder_object,
  ggplot(data = curve) +
    geom_line(aes(x = rank, y = ES), color = "green") +
    # geom_ribbon(
    #   data = stats,
    #   mapping = aes(
    #     x = rank, ymin = 0,
    #     ymax = stat / maxAbsStat * (spreadES / 4)
    #   ),
    #   fill = "grey"
    # ) +
    geom_segment(
      data = ticks,
      mapping = aes(
        x = rank, y = -spreadES / 16,
        xend = rank, yend = spreadES / 16
      ),
      size = 0.2
    ) +
    geom_hline(yintercept = posES, colour = "red", linetype = "dashed") +
    geom_hline(yintercept = negES, colour = "red", linetype = "dashed") +
    geom_hline(yintercept = 0, colour = "black") +
    theme(
      panel.background = element_blank(),
      panel.grid.major = element_line(color = "grey92")
    ) +
    labs(x = "rank", y = "enrichment score") +
    geom_text_repel(
      mapping = aes(x = rank, y = ES, label = gene_symbol), data = edge %>% filter(rank %in% ticks$rank) %>% make_selection(),
      max.overlaps = Inf, max.time = 4
    )
  # geom_text_repel( mapping = aes(x = rank, y = ES, label = gene_symbol), data = edge %>% filter(rank %in% ticks$rank) )
)
```

```{r, eval=F}
rankorder_object <- fgsea_tools$get_rankorder(genesets_of_interest[[1]], rankobject)

enplot_data <- plotEnrichmentData(genesets_of_interest[[1]], rankobject)

.for_merge <- list_of_geneset_dfs[[collection_name]] %>%
  filter(gs_name == geneset_of_interest_name) %>%
  mutate(entrez_gene = as.character(entrez_gene)) %>%
  distinct(entrez_gene, .keep_all = T) # sometimes there's a duplicate annotate (maybe because multiple ensembl gids)


.rankorder <- rankorder_object$edge %>% left_join(
  .for_merge,
  by = c("id" = "entrez_gene")
)
rankorder_object$edge <- .rankorder # %>% filter(!is.na(stat_tick))
rankorder_object$stats <- .rankorder # %>% filter(!is.na(stat_tick))

rankorder_object$edge %>%
  make_selection() %>%
  dim()

plot_tools$edgeplot1(rankorder_object = rankorder_object) +
  geom_text_repel(data = . %>% filter(!is.na(stat_tick)) %>% make_selection(), aes(label = gene_symbol), max.overlaps = Inf)
```

```{r, eval=F}
plot_tools$edgeplot1(rankorder_object = rankorder_object) +
  geom_text_repel(data = . %>% filter(!is.na(stat_tick)) %>% make_selection(), aes(label = gene_symbol), max.overlaps = Inf)
```

```{r, fig.width=10, fig.height=8, eval=F}
make_selection <- function(x) {
  es_min <- x %>%
    pull(ES) %>%
    min(na.rm = T)
  es_max <- x %>%
    pull(ES) %>%
    max(na.rm = T)
  .first <- which(x$ES == es_min)
  .second <- which(x$ES == es_max)
  .top <- x %>%
    arrange(rank) %>%
    head(.first)
  .bot <- x %>%
    arrange(rank) %>%
    tail(dim(x)[1] - .second)
  return(bind_rows(.top, .bot))
}
```

```{r, fig.width=10, fig.height=8, eval=F}
rankorder %<>% arrange(stat_tick)
rankorder %>%
  plot_tools$edgeplot1() +
  geom_text_repel(data = . %>% make_selection(), aes(label = gene_symbol), max.overlaps = Inf)
```



```{r, fig.width=12, fig.height=10, eval=F}
# rankorder_all <- .GlobalEnv$get_rankorder(ranks_list[[1]], pathways_list_of_lists$H_$HALLMARK_ADIPOGENESIS  )

rankorder_all <- fgsea_tools$get_rankorder(ranks_list[[1]], pathways_list_of_lists$H_$HALLMARK_ADIPOGENESIS)

rankorder_all2 <- rankorder_all %>%
  left_join(
    list_of_geneset_dfs$H_ %>%
      mutate(entrez_gene = as.character(entrez_gene)) %>%
      distinct(entrez_gene, .keep_all = T),
    by = c("id" = "entrez_gene")
  )



pdata <- p$data
pdata_more <- left_join(
  pdata,
  rankorder_all2
) %>% arrange(rank)


make_selection2 <- function(x) {
  .top <- x %>%
    arrange(rank) %>%
    filter(!is.na(ES) & !is.na(gene_symbol)) %>%
    head(10)
  .bot <- x %>%
    arrange(rank) %>%
    filter(!is.na(ES) & !is.na(gene_symbol)) %>%
    tail(10)
  return(bind_rows(.top, .bot))
}


p <- fgsea::plotEnrichment(pathways_list_of_lists$H_$HALLMARK_ADIPOGENESIS, ranks_list[[1]])
p + geom_text_repel(data = rankorder %>% make_selection(), aes(x = rank, y = ES, label = gene_symbol), max.overlaps = Inf)
# print(p)
```
